Poster figures

### 2025-02-27 analysis/poster_ncar/presentation_figures.Rmd
More figures for ncar

@@

```{r fig-hindcast-extent}
#| fig-cap: Median sea ice extent for al hindcasts initialised the first of the month for +S2 and +S1 in colours representing the start month. In black, the median sea ice extent of +CDR. 
hindcast_extent |> 
  _[, median_ci(aice), by = .(forecast_time = update(forecast_time, year = 2001),
                              lag = as.numeric(as.Date(time) - forecast_time),
                              model)] |>
  _[, time := forecast_time + lag] |>
  _[, time2 := update(time, year = 2001)] |>
  _[, month := lubridate::month(forecast_time, label = TRUE)] |>
  # _[]
  ggplot(aes(time2, mid)) +
  geom_line(data = extent_climatology[dataset == which_dataset], 
            aes(x = as.Date(time),
                group = dataset)) + 
  geom_line(aes(color = month, group = interaction(forecast_time, time != time2))) +

@@

```{r}
hindcast_extent |> 
  _[, median_ci(aice), by = .(forecast_time = update(forecast_time, year = 2001),
                              lag = as.numeric(as.Date(time) - forecast_time),
                              model)] |>
  _[, time := forecast_time + lag] |>
  _[, time2 := update(time, year = 2001)] |>
  _[, month := lubridate::month(forecast_time, label = TRUE)] |>
  _[copy(extent_climatology)[, time := as.Date(time)], on = c(time2 = "time")] |> 
  _[] |> 
  ggplot(aes(time2, mid - i.mid)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(color = month, group = interaction(forecast_time, time != time2))) +

@@

```{r hindcast_clim}
zeropad <- function(x, n = 2) formatC(x, width = 2, flag = "0")
monnb <- function(d) { lt <- as.POSIXlt(as.Date(d, origin="1900-01-01")); 
lt$year*12 + lt$mon } 
mondf <- function(d1, d2) { monnb(d2) - monnb(d1) }

@@

hindcast_clim <- lapply(setNames(months, months), \(month) {
  here::here(glue::glue("data/derived/climatology/S2/{zeropad(month)}/em.nc")) |> 
    cdo_selmonth(month) |> 
    cdo_execute() |> 
    ReadNetCDF("aice")
}) |> 
  rbindlist(idcol = "forecast_month") |> 
  _[, forecast_month := as.numeric(forecast_month)]

hindcast_clim <- hindcast_clim |> 
  copy() |> 
  _[, .(aice = mean(aice)), by = .(month(time), ygrid, xgrid, forecast_month)] |> 
  _[forecast_month == month]

obs_clim <- CDR() |> 
  climatology() |> 
  cdo_ymonmean() |>
  cdo_execute() |> 
  ReadNetCDF(c(obs = "aice")) |> 
  _[, month := month(time)]

hindcast_clim <- hindcast_clim |> 
  merge(obs_clim)
```

```{r}
trans_pink <- gglgbtq::palette_lgbtq("transgender")[2] |> 
  colorspace::darken(0.2)
trans_blue <- gglgbtq::palette_lgbtq("transgender")[1] |> 
  colorspace::darken(0.2)

@@

```{r fig-bias}
#| fig-width: 9
#| fig-height: 6
#| fig-cap: Mean sea ice concentrations 1-month lag monthly mean +S2 forecast bias compared with NSIDC.

hindcast_clim |> 
  # _[month == 1] |>
  # dcast(time + xgrid + ygrid ~ model, value.var = "aice") |> 
  # setnames(which_dataset, "obs") |>
  ggplot(aes(xgrid, ygrid)) +
  geom_contour_fill(aes(z = aice - obs, fill = after_stat(level)), 
                    breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)) +
  # geom_contour2(aes(z = aice - obs), breaks = AnchorBreaks(binwidth = 0.1, exclude = 0),
  #               linewidth = 0.1) +
  scale_fill_divergent_discretised(paste0("Bias with respect to ",
                                          labels_models[which_dataset]),
                                   low = trans_pink,
                                   high = trans_blue) +
  geomcoord_antarctica +
  geom_antarctica_fill +
  facet_wrap(~ month, 
             labeller = labeller(month = setNames(months, month.abb)),
             nrow = 2) +
  wide_legend +
  tag_facets()
```

@@

scale_fill_models <-   ggplot2::scale_fill_manual(NULL,
                                                  values = colours_models,
                                                  labels = labels_models
)
```

```{r fig-extent-anom}
#| fig-cap: Sea ice extent anomalies for +S1 and +S2 (black) and +CDR (blue).

@@

```{r}
g %+% g$data[version != "climatology"]
```

```{r}
g %+% g$data[version != "persistence"]
```

```{r}
p <- g

```{r}
p <- g +
  coord_cartesian(xlim = c(0, 100))

>>>>

scale_fill_models <-   ggplot2::scale_fill_manual(NULL,
                                                  values = colours_models,
                                                  labels = labels_models
)

trans_pink <- gglgbtq::palette_lgbtq("transgender")[2] |> 
  colorspace::darken(0.2)
trans_blue <- gglgbtq::palette_lgbtq("transgender")[1] |> 
  colorspace::darken(0.2)

@@

```{r}
cm <- 2.54
```

```{r fig-hindcast-extent, fig.height=11.99/cm, fig.width=23.59/cm}
#| fig-cap: Median sea ice extent for al hindcasts initialised the first of the month for +S2 and +S1 in colours representing the start month. In black, the median sea ice extent of +CDR. 
hindcast_extent |> 
  _[, median_ci(aice), by = .(forecast_time = update(forecast_time, year = 2001),
                              lag = as.numeric(as.Date(time) - forecast_time),
                              model)] |>
  _[, time := forecast_time + lag] |>
  _[, time2 := update(time, year = 2001)] |>
  _[, month := lubridate::month(forecast_time, label = TRUE)] |>
  # _[]
  ggplot(aes(time2, mid)) +
  geom_line(data = extent_climatology[dataset == which_dataset], 
            aes(x = as.Date(time),
                group = dataset)) + 
  geom_line(aes(color = month, group = interaction(forecast_time, time != time2))) +

@@

```{r hindcast_extent-bias, fig.height=11.99/cm, fig.width=23.59/cm}
hindcast_extent |> 
  _[, median_ci(aice), by = .(forecast_time = update(forecast_time, year = 2001),
                              lag = as.numeric(as.Date(time) - forecast_time),
                              model)] |>
  _[, time := forecast_time + lag] |>
  _[, time2 := update(time, year = 2001)] |>
  _[, month := lubridate::month(forecast_time, label = TRUE)] |>
  _[copy(extent_climatology)[, time := as.Date(time)], on = c(time2 = "time")] |> 
  _[] |> 
  ggplot(aes(time2, mid - i.mid)) +
  geom_hline(yintercept = 0) +
  geom_line(aes(color = month, group = interaction(forecast_time, time != time2))) +

@@

```{r extent-delta, fig.height=11.99/cm, fig.width=19.39/cm}

dif_hindcast <- hindcast_extent |> 
  copy() |> 
  _[order(time)] |> 
  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(model, forecast_time, member)] |> 
  _[, median_ci(dif), by = .(model, time = update(time, year = 2000))]

extent_daily |> 
  _[order(time)] |> 
  _[, dif := c(NA, diff(aice)/diff(as.numeric(as.Date(time)))), by = .(dataset)] |> 
  _[, median_ci(dif), by = .(dataset, time = update(time, year = 2000))] |> 
  _[dataset == which_dataset] |> 
  ggplot(aes(time, mid)) +
  geom_hline(yintercept = 0, color = "gray50", linewidth = 0.2) +
  geom_smooth(aes(group = dataset, color = dataset),
              span = 30/366, linewidth = 0.5, se = FALSE, n = 366)  +
  geom_smooth(data = dif_hindcast,
              aes(color = model), span = 30/366, linewidth = 0.5, se = FALSE, n = 366) +
  scale_x_datetime(NULL, date_breaks = "1 month", 
                   date_labels = "%b", expand = c(0, 0)) +
  scale_y_continuous(NULL, labels = \(x) labels_extent(x, units = "M km²/day")) +
  scale_color_models
```

```{r hindcast_clim}
zeropad <- function(x, n = 2) formatC(x, width = 2, flag = "0")
monnb <- function(d) { lt <- as.POSIXlt(as.Date(d, origin="1900-01-01")); 
lt$year*12 + lt$mon } 
month_differnece <- function(d1, d2) { monnb(d2) - monnb(d1) }

month_diff <- function(month1, month2) {
  d <- (month2 - month1)

  (d %% 12)

}

@@

hindcast_clim <- purrr::pmap(CJ(month = 1:12, version = c("S2", "S1")), \(month, version) {
  here::here(glue::glue("data/derived/climatology/{version}/{zeropad(month)}/em.nc")) |> 
    # cdo_selmonth(month) |> 
    cdo_monmean() |> 
    cdo_execute() |> 
    ReadNetCDF("aice") |> 
    _[, let(forecast_month = month, 
            version = version)] 
}) |> 
  rbindlist() |> 
  _[, forecast_month := as.numeric(forecast_month)] |> 
  _[, forecast_time := update(time, month = forecast_month)] |> 
  _[, time := as.Date(update(time, day = 1))]

obs_clim <- CDR() |> 
  climatology() |> 
  cdo_ymonmean() |>
  cdo_execute() |> 
  ReadNetCDF(c(obs = "aice")) |> 
  _[, time := as.Date(update(time, day = 1))]

hindcast_clim <- hindcast_clim |> 
  merge(obs_clim) |> 
  _[, month := month(time)] |>
  _[, month_diff := month_differnece(time[1], forecast_time[1]), by = .(forecast_month, time)]

hindcast_clim[, month_diff := month_diff(forecast_month[1], month[1]), by = .(forecast_month, month)]

@@

```{r fig-bias}
#| fig-cap: Mean sea ice concentrations 1-month lag monthly mean +S2 forecast bias compared with NSIDC.

bias_plot <- function(hindcast_clim, which_version, which_lead) {
  hindcast_clim |> 
    _[version == which_version] |>
    _[month_diff == which_lead] |> 
    # dcast(time + xgrid + ygrid ~ model, value.var = "aice") |> 
    # setnames(which_dataset, "obs") |>
    ggplot(aes(xgrid, ygrid)) +
    geom_contour_fill(aes(z = aice - obs, fill = after_stat(level)), 
                      breaks = AnchorBreaks(binwidth = 0.1, exclude = 0)) +
    # geom_contour2(aes(z = aice - obs), breaks = AnchorBreaks(binwidth = 0.1, exclude = 0),
    #               linewidth = 0.1) +
    scale_fill_divergent_discretised(paste0(which_version, " bias at leadtime ", which_lead, " months"),
                                     low = trans_pink,
                                     high = trans_blue) +
    geomcoord_antarctica +
    geom_antarctica_fill +
    facet_wrap(~ month, 
               labeller = labeller(month = setNames(months, month.abb)),
               nrow = 2) +
    wide_legend +
    tag_facets()
}

```{r, fig.width=22.15/cm, fig.height=12.3/cm}
hindcast_clim |> 
  bias_plot("S2", 0)
```

```{r}
hindcast_clim |> 
  bias_plot("S1", 0)
```

```{r}
hindcast_clim |> 
  bias_plot("S2", 6)

```{r}
hindcast_clim |> 
  bias_plot("S1", 6)
```

@@

```{r fig-extent-anom, fig.height=11.99/cm, fig.width=23.22/cm}
#| fig-cap: Sea ice extent anomalies for +S1 and +S2 (black) and +CDR (blue).

@@

```{rrmse_persistence, fig.width=19.43/cm, fig.height=11.99/cm}
g %+% g$data[version != "climatology"]
```

```{rrmse_climatology, fig.width=19.43/cm, fig.height=11.99/cm}
g %+% g$data[version != "persistence"]
```

```{r}
# ACHTUNG: we have something going on with the 29th of february.

clim_std2 |> 
  rbind(errors) |> 
  _[measure == "rmse"] |> 
  _[obs_dataset == which_dataset] |> 
  _[, .(rmse = median(value, na.rm = TRUE)), 
    by = .(time = update(time, year = 2001), 
           time_forecast = update(time_forecast, year = 2001), 
           lag,
           version)] |> 
  # _[version == "climatology"] |>
  ggplot(aes(time, rmse)) +
  geom_line(data = \(x) x[version == "climatology"],
            aes(group = interaction(time_forecast, month(time_forecast) <= month(time)))) +
  # geom_line(data = \(x) x[version == "S2"],
  #           aes(color = factor(month(time_forecast)),
  #               group = interaction(time_forecast, month(time_forecast) <= month(time)))) +
  # geom_point(data = \(x) x[version == "S2" & lag ==1], aes(color =  factor(month(time_forecast)))) +
  scale_color_manual(values = colorRampPalette(gglgbtq::palette_lgbtq("nonbinary")[c(1, 3, 4, 1)])(12),
                     aesthetics = c("fill", "color"), guide = "none") +
  scale_y_continuous(limits = c(0, NA)) +
  geom_vline(xintercept = as.Date("2001-02-28"))
```

```{r, fig.width=19.43/cm, fig.height=11.99/cm}
p <- g +
  coord_cartesian(xlim = c(0, 100))
